---
title: "Computational Musicology"
author: "Xing Yu Pan"
date: "spring 2021"
output: html_document
---


## *** PLEASE DO NOT SHOW THIS PORTFOLIO IN CLASS ***
#### Overview Corpus 

For this portfolio I choose to compare two of my favorite composers: Joe Hisaishi and Hans Zimmer. They are both know for their scores in the film industry with their own unique styles. I'm interested in how they differ in the way they compose and how this can be explained by the geographic region, as Joe Hisaishi mostly works with Japanese films and Hans Zimmer with his career in Hollywood. However it is not yet clear which part I will be compare to the most. 
  
#### Short Introduction of the composers

As Joe Hisaishi is one of the most famous Japanese composer, known for over 100 film scores and the association with the animator Hayao Miyazaki, he has provide many composition which has worldwide success, including the opening theme of Miyazaki's film Spirited Away and later on the score of Howl's Moving Castle. His composingstyle can be described as gentle, minimalistic and melancholic, which also becomes a tradematk of much of the studio Ghibli's output. 

Hans Zimmer is a German film score composer and producer. His work are notable for integreting electronic sounds with traditional orchestral arrangements. He has many award winning compositions include The Lion King and the Pirates of the Caribbean series. 

#### Some basic comparisons 

```{r setup, include=FALSE,  warning=FALSE}

library(tidyverse)
library(spotifyr)
```
```{r, echo=FALSE, warning=FALSE }
Joe <- get_artist_audio_features("7nzSoJISlVJsn7O0yTeMOB")
Hans <- get_artist_audio_features("0YC192cP3KPCRWx8zr8MfZ")

```
```{r, echo=FALSE,  warning=FALSE}
awards <-
  bind_rows(
    Joe %>% mutate(category = "Joe"),
    Hans %>% mutate(category = "Hans")
  )

```
```{r, echo=FALSE,  warning=FALSE}
awards %>%
  ggplot(aes(x = category, y = valence, fill = category)) +
   geom_boxplot() + theme_get() + ylab("Overall energy") + 
           xlab("Category")

```


As shown in the graph above, both artists have a relatively low energy in terms of intensity and activity of their audio features. The graph indicates that Joe Hisaishi has a larger variance in energy compared to Hans Zimmer, however the average mean of their energy level is pretty similar. To have a more understanding in the meaning of the feature 'energy', we introduce another feature 'valence'. In the graph below valence is plotted against the energy of both artists. Each point indicates an album of artist. As we can see from the scatterplot, the density is a largest in the left-under area where both energy and valence is low. Hans Zimmer's work mostly has a very low valence level(0-0.125) and a low to medium energy level(0-0.50), while Joe Hisaishi's work has a more even contribution across the left-under section. 
  From both graph we can conclude that both artist have a pretty similar composing style where the valence and the energy is low, i.e. the soundtracks sound pretty sad according to the music theory (- valence, - energy). 



```{r, echo=FALSE,  warning=FALSE}
p <- ggplot(awards, aes(x = valence, y = energy, color = category)) + geom_point(position = "jitter") + scale_x_continuous("valence") +  scale_color_discrete("Artist")

p + geom_hline(yintercept=0.50, linetype="dashed", color = "red") + geom_vline(xintercept = 0.5,  linetype="dashed", color = "red")

```





```


